Debiased Calibration for Generalized Two-Phase Sampling

### 2024-04-27 final/reports/prelim_paper_ov.tex
Updated Topic 1 simulation results. We are using a SRS-Poisson setup.

\begin{theorem}[Design Consistency]
  Under some regularity conditions,

@@

  where
\textcolor{red}{
   \begin{eqnarray*}
   \hat Y_{\ell}( \lambda^*, \phi^*) &=&   \hat{Y}_{\pi^*} + 
  \left(\sum_{i \in A_1} w_{1i} x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1} x_i  \right)^T \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i   -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
  \end{eqnarray*} 
  and  $$\begin{pmatrix}
  \phi_1^* \\
  \phi_2^* 
  \end{pmatrix}
  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
  \begin{pmatrix}
  \bx_i \bx_i^T &   \bx_i g_i   \\
  g_i  \bx_i^T   & g_i^2     \end{pmatrix} \right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} \begin{pmatrix}
    \bx_i \\ g_i 
  \end{pmatrix}y_i $$
with $g_i = g( \pi_{2i |1}^{-1}) q_i$. }
\end{theorem}

\textcolor{red}{I have tried to improve the expression of the linearization formula. Please check if my algebra above is correct. }
\begin{proof}
  In this proof, we derive the solution to Equation~\ref{eq:primal} and show
  that it is asymptotically equivalent to a regression estimator. Using the
  method of Lagrange multipliers, to solve Equation~\ref{eq:primal} we need to
  minimize,

  $$\mathcal{L}(w_{2|1}) = \sum_{i \in A_2} w_{1i} G(w_{2i|1}) q_i + 
  \lambda \left(\sum_{i \in A_2} w_{1i} w_{2i|1} z_i q_i - \sum_{i \in A_1} w_{1i}
  z_i q_i\right).$$

  The first order conditions show that

  $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i +
    \lambda w_{1i} z_i q_i = 0.$$

  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda z_i)$ and $\hat \lambda$ is
  determined by the calibration condition. When the sample size gets large, we
  have $\hat w_{2i|1}(\hat \lambda) \to d_{2i|1}$ which means that $\hat \lambda
  \to \lambda^*$ where $\lambda^* = (\bf 0, 1)$. Then using the linearization
  technique of \cite{randles1982asymptotic}, we can construct a regression
  estimator,

@@

  \begin{align*}
    \hat Y_{DCE}(\hat \lambda) 
    &= \hat Y_\ell(\lambda^*, \phi^*) + 
    E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*,
    \phi^*)\right](\hat \lambda - \lambda^*) + \frac{1}{2}
    E\left[\frac{\partial}{\partial \lambda^2} \hat Y_{DCE}(\lambda^*)\right] (\hat
    \lambda - \lambda^*)^2\\
    &= \hat Y_\ell(\lambda^*, \phi^*) + O(N)o_p(n_2^{-1}).
  \end{align*}

  The final equality comes from the fact that 
  $E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*, \phi^*)\right]
  = 0$, $\frac{\partial}{\partial \lambda^2} \hat w_{2i|1}(\lambda^*)$ is
  bounded and $|\hat \lambda - \lambda^*| = o_p(n_2^{-1/2})$, which proves our
  result.
\footnote{$\hat{\lambda}$ should satisfy $\hat{\lambda} - \lambda^* = O_p (n_2^{-1/2})$. }
\end{proof}

%\begin{proof}
%  First, we derive the solution for $\hat Y_{DCE}$, and then derive the
%  asymptotic variance of this estimator using the linearization technique of
%  \cite{randles1982asymptotic}.
%
%  We want to solve Equation~\ref{eq:primalloss} subject to the constraints in
%  Equations~\ref{eq:calconst1} and \ref{eq:calconst2}. Using the method of
%  Lagrange multipliers we have a Lagrangian of 
%
%  $$\mathcal{L}(w_2) = -\sum_{i \in A_2} w_{1i} G(w_{2i})q_i + \lambda 
%  \left(\sum_{i \in A_2} w_{1i} w_{2i} z_i - \sum_{i \in A_1} w_{1i} z_i\right).$$
%
%  Our first order conditions yield,
%
%  \begin{align*}
%    0 &= \frac{\partial}{\partial w_{2i}} \mathcal{L}(w_{2}) \\ 
%      &= -w_{1i} g(w_{2i}) q_i + \lambda w_{1i} z_i.
%  \end{align*}
%
%  This means that $\hat w_{2i} = g^{-1}(\hat \lambda z_i / q_i)$ and since $\hat
%  \lambda$ is determined from the solution to Equations~\ref{eq:calconst1} and
%  \ref{eq:calconst2}. Hence, $\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} 
%  \hat w_{2i}(\hat \lambda) y_i$.
%
%  To get the variance of $\hat Y_{DCE}$ we use the linearization technique of
%  \cite{randles1982asymptotic}. First, notice that 
%
%  $$\hat Y_{DCE} = \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i =
%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) y_i + \left(
%  \sum_{i \in A_1} w_{1i} z_i -
%  \sum_{i \in A_2} w_{1i} \hat w_{2i}(\hat \lambda) z_i\right)\phi
%  := \hat Y_\ell(\hat \lambda, \phi).$$
%
%  We can find the value of $\phi$ such that $E[\frac{\partial}{\partial
%  \lambda}\hat Y_\ell(\lambda, \phi)]_{\lambda = \lambda^*}$ where $\lambda^*$
%  is the probability limit of $\hat \lambda$. Notice, that 
%  $\lambda^* = (\bf 0, 1)$ because $\hat w_{2i} \to \pi_{2i|1}^{-1}$. This means
%  that the optimal value of $\phi$ solves,
%
%  $$E\left[\frac{\partial}{\partial \lambda} \hat Y_\ell(\lambda, \phi)\right] =
%  E\left[\sum_{i \in A_2} w_{1i} \frac{\partial}{\partial \lambda} w_{2i|1}(\lambda)
%    y_i - \phi \sum_{i \in A_2} w_{1i} z_i \frac{\partial}{\partial \lambda}
%    w_{2i|1}(\lambda) / q_i \right]$$
%
%  which means that
%
%  $$\phi^* = \left[\sum_{i \in U} \pi_{2i|1} \left(\frac{z_i z_i^T}{
%    g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in U} \pi_{2i|1}
%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
%
%  Hence,
%
%  \begin{align*}
%    \hat Y_\ell(\lambda^*, \phi^*) 
%    &= \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1}
%    y_i + \left(\sum_{i \in A_1} w_{1i} z_i / q_i -
%      \sum_{i \in A_2} w_{1i} \pi_{2i|1}^{-1} z_i / q_i\right)\phi^* \\ 
%    &= \sum_{i \in A_1} w_{1i} \eta_i
%  \end{align*}
%
%  where 
%
%  $$\eta_i = z_i \phi^* / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
%  - z_i \phi^* / q_i\right).$$
%
%  To analyze the variance, consider the reverse framework of
%  \cite{fay1992inferences}, \cite{kim2006replication}, and \cite{shao1999variance}.
%  In this setup, we assume that $\mathcal{R} = \{\delta_{2i|1}\}_{i \in U}$ is
%  known at the
%  finite population level. Then with this we have,
%
%  \begin{align*}
%    V(\hat Y_{\ell}(\lambda^*, \phi^*)) 
%    &= V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\right) \\
%    &= V\left(E\left[\sum_{i \in A_1} w_{1i}\eta_i(\phi^*)\mid\mathcal{R}\right]\right) 
%    +E\left[V\left(\sum_{i \in A_1} w_{1i} \eta_i(\phi^*)\mid\mathcal{R}\right)\right]\\
%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right) + E\left[\sum_{i \in U} 
%    \sum_{j \in U} \Cov(\delta_{1i}, \delta_{1j}) 
%    w_{1i} w_{1j} \eta_i(\phi^*) \eta_j(\phi^*)\right]\\
%  \end{align*}
%
%  We can estimate this variance using 
%
%  $$\hat V_{DCE} = \sum_{i \in A_1} \sum_{j \in A_1} \left(\frac{\pi_{1ij} -
%  \pi_{1i}\pi_{1j}}{\pi_{1ij}}
%    w_{1i} w_{1j} \hat \eta_i \hat \eta_j \right)
%  $$
%
%  for $\hat \eta_i = z_i \hat \phi / q_i + \frac{\delta_{2i|1}}{\pi_{2i|1}} \left(y_i
%  - z_i \hat \phi / q_i\right)$ where 
%
%  $$\hat \phi = \left[\sum_{i \in A_2} w_{1i}\left(\frac{z_i z_i^T}{
%  g'(g^{-1}(g(d_{2i|1}) / q_i))q_i^2}\right)\right]^{-1} \sum_{i \in A_2} w_{1i}
%    \left(\frac{z_i y_i}{g'(g^{-1}(g(d_{2i|1}) / q_i)) q_i}\right).$$
%
%  The problem with this estimator is that it is biased as $E[\hat V_{DCE}]$ is
%  asymptotically equivalent to
%  $V\left(\sum_{i \in A_1} w_{1i} \eta(\phi^*)\right)$. This yields a bias of 
%
%  \begin{align*}
%    \text{Bias}(\hat V_{DCE}) 
%    &= V\left(\sum_{i \in U} \eta_i(\phi^*)\right)\\ 
%    &= \sum_{i \in U} \sum_{j \in U} \Delta_{2ij|1} w_{2i|1}
%    w_{2j|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).
%  \end{align*}
%
%  where $\Delta_{2ij|1} = \pi_{2ij|1} - \pi_{2i|1} \pi_{2j|1}$. 
%  However, we can estimate this bias with
%
%  $$\widehat{\text{Bias}}(\hat V_{DCE}) = 
%  \sum_{i \in A_2} \sum_{j \in A_2} w_{1ij} w_{2ij|1} w_{2i|1} w_{2j|1} 
%  \Delta_{2ij|1} (y_i - z_i \phi^* / q_i)(y_j - z_j \phi^* / q_j).$$
%
%\end{proof}

\subsection{Simulation Studies}

@@

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
X_{4i} &\stackrel{ind}{\sim} Unif(0.1, 0.9) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
\pi_{1i} &= \Phi_3(-x_{3i} - 2) \\
\pi_{2i|1} &= x_{4i}.
\end{aligned}
$$

where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.

\textcolor{red}{The simulation setup is somewhat strange. The first order inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$. In this case, your second-phase sampling is essentially equivalent to the simple random sampling. }
This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
consider a finite population of size $N = 10,000$ with both the Phase 1 and
Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
Phase 1 sample
size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
$E[n_2] \approx 550$. In the Phase 1 sample, we observe 
$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
simulation does not deal with model misspecification, and we compare the
proposed method for the parameter $\bar Y_N$ with four approaches:

\begin{itemize}
  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
  \item[2.] Two Phase Regression estimator (TP-Reg): 
    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
    where $\hat \beta = 
    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
    and $\bf x_i = (x_{1i}, x_{2i})^T$,
    \textcolor{red}{Dr. Kim, should I modify the simulation so that the
    regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This is
    the method from \cite{kwon2024debiased} with the true population level
    constraints, and 
  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
    This is the proposed method with the Phase 1 sample being used to estimate
    the population level constraints.
\end{itemize}

In addition to estimating the mean parameter $\bar Y_N$, we also construct
variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
approach we have the following variance estimate\footnote{These variance
estimates use the fact that we have Poisson sampling for both phases in the
simulation.}:

\begin{figure}[ht!]
  \centering
  \begin{tabular}{lcc}
    \toprule
    Estimator & Variance & Notes \\
    \midrule
    $\pi^*$ & $N^{-2} \sum_{i \in A_2} \left(\pi_{2i|1}^{-2} - \pi_{2i|1}^{-1}\right)
    y_i^2 $ & \\
    TP-Reg  & 
      {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} \left(\pi_{1i}^{-2} -
          \pi_{1i}^{-1}\right) 
      \eta_i^2 
      + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(\pi_{2i|1}^{-1}
  - 1) (y_i - \bf x_i' \hat \beta)^2 \right)$}
    & 
      {\scriptsize$\eta_i = \bf x_i \hat \beta + \frac{\delta_{2i}}{\pi_{2i|1}}(y_i - 
      \bf x_i \hat \beta)$} \\
      %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
      DC-Pop  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma)$ & 
      {\scriptsize$\Pi
      = \text{diag}(1 - (\pi_{1}\pi_{2|1})^{-1}) \cdot \frac{\hat w^2}{N^2}$} \\
      DC-Est  & $(Y - \bf z^T \hat \gamma)^T \Pi (Y - \bf z^T \hat \gamma) +
      \hat \gamma_{[1:3]}^T V(\bf x) \hat \gamma_{[1:3]}/N^2$ & \\
    \bottomrule
  \end{tabular}
  \caption{This table gives the formulas for each variance estimator used in
  this simulation.}
  \label{tab:varforms}
\end{figure}

We run this simulation $1000$ times for each of these methods and compute the
Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
unbiasedness of each estimator. The results are in Figure~\ref{fig:tpdc-mean}.

\begin{figure}[ht!]
  \centering
\input{tables/tpdcsim_mean.tex}
% This table was generated from /src/explore/20240411-tpdcsim.qmd
\caption{This table shows the results of the simulation study. It displays the
Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
unbiasedness of each estimator for the estimators: $\pi^*$, TP-Reg, DC-Pop, and
DC-Est.}
\label{fig:tpdc-mean}
\end{figure}

@@

$$\hat X_1 = \sum_{i \in A_1}^{n_1} \frac{x_i}{\pi_{1i}} \text{ and } 
\hat X_2 = \sum_{i \in A_2}^{n_2} \frac{x_i}{\pi_{2i}}$$

where $\pi_{1i}$ is the probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in
A_2)$. \textcolor{red}{I don't know if it makes sense to change the Phase 2
selection probability as $\pi_{2i|1}$. Should I keep it at $\pi_{2i}$? But if
I do, how is the two phase setup useful?} We can combine these estimates to get
$\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
\cite{merkouris2004combining} for the optimal choice of $W$). We can then define
a regression estimator as

$$
\hat Y_{NN, reg} = \hat Y_2 + (\hat X_c - \hat X_2)^T \hat \beta_q = 
\hat Y_2 + (\hat X_1 - \hat X_2)^T W\hat \beta_q 
$$

$$\hat \beta_q = \left(\sum_{i \in A_2} \frac{x_i x_i^T}{q_i}\right)^{-1}
\sum_{i \in A_2} \frac{x_i y_i}{q_i} \text{ and }\hat Y_2 = \sum_{i \in A_2}
\frac{y_i}{\pi_{2i}}.$$

@@

>>>>

% \textcolor{red}{In Theorem 1, you used $\hat{\lambda}$ without defining it.  }
\begin{theorem}[Design Consistency]\label{thm:dc1}
  Let $\lambda^*$ be the probability limit of $\hat \lambda$.
  Under some regularity conditions,

@@

  $$\hat Y_{\ell}(\hat \lambda, \phi^*) = \hat Y_{DCE}(\hat \lambda) + 
  \left(\sum_{i \in A_1} w_{1i} z_i q_i - \sum_{i \in A_2} w_{1i} \hat w_{2i|1}(\hat
  \lambda) z_i q_i\right)\phi^*$$

  $$\phi^* = 
  \left[\sum_{i \in U} \frac{\pi_{2i|1}z_i z_i^T q_i}{g'(d_{2i|1})}\right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}z_i y_i}{g'(d_{2i|1})}.$$

   \begin{eqnarray*}
   \hat Y_{\ell}( \lambda^*, \phi^*) &=&   \hat{Y}_{\pi^*} + 
  \left(\sum_{i \in A_1} w_{1i} x_i  -  \sum_{i \in A_2} w_{1i} \pi_{2i \mid
  1}^{-1} x_i  \right)^T \phi_1^* + \left(\sum_{i \in A_1} w_{1i} g_i   -
\sum_{i \in A_2} w_{1i} \pi_{2i \mid 1}^{-1}g_i  \right)^T \phi_2^*  
  \end{eqnarray*} 
  and  $$\begin{pmatrix}
  \phi_1^* \\
  \phi_2^* 
  \end{pmatrix}
  = \left[ \sum_{i \in U} \frac{\pi_{2i \mid 1} }{ g'(d_{2i|1}) q_i} 
  \begin{pmatrix}
  \bx_i \bx_i^T &   \bx_i g_i   \\
  g_i  \bx_i^T   & g_i^2     \end{pmatrix} \right]^{-1}
  \sum_{i \in U} \frac{\pi_{2i|1}}{ g'(d_{2i | 1}) q_i} \begin{pmatrix}
    \bx_i \\ g_i 
  \end{pmatrix}y_i $$
with $g_i = g( \pi_{2i |1}^{-1}) q_i$.
\end{theorem}

\begin{proof}
  In this proof, we derive the solution to Equation~\ref{eq:primal} and show
  that it is asymptotically equivalent to a regression estimator. Using the
  method of Lagrange multipliers, to solve Equation~\eqref{eq:primal} we need to
  minimize the Lagrangian in Equation~\eqref{eq:legragedc1}. 
  The first order conditions show that

  $$\frac{\partial \mathcal{L}}{\partial w_{2i|1}}: g(w_{2i|1}) w_{1i}q_i -
    \lambda w_{1i} z_i q_i = 0.$$

  Hence, $\hat w_{2i}(\lambda) = g^{-1}(\lambda^T z_i)$ and $\hat \lambda$ is
  determined by Equation~\eqref{eq:lamdc1}. When the sample size gets large, we
  have $\hat w_{2i|1}(\hat \lambda) \to d_{2i|1}$ which means that $\hat \lambda
  \to \lambda^*$ where $\lambda^* = (\bf 0, 1)$. Then using the linearization
  technique of \cite{randles1982asymptotic}, we can construct a regression
  estimator,

@@

  \begin{align*}
    \hat Y_{DCE}(\hat \lambda) 
    &= \hat Y_\ell(\lambda^*, \phi^*) + 
    E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*,
    \phi^*)\right](\hat \lambda - \lambda^*) + \frac{1}{2}
    E\left[\frac{\partial}{\partial \lambda^2} \hat Y_{DCE}(\lambda^*)\right] (\hat
    \lambda - \lambda^*)^2\\
    &= \hat Y_\ell(\lambda^*, \phi^*) + O(N)O_p(n_2^{-1}).
  \end{align*}
%\textcolor{red}{(The remainder term shoould be $O_p(N)O_p(n_2^{-1})$.)}
  The final equality comes from the fact that 
  $E\left[\frac{\partial}{\partial \lambda}\hat Y_\ell(\lambda^*, \phi^*)\right]
  = 0$, $\frac{\partial}{\partial \lambda^2} \hat w_{2i|1}(\lambda^*)$ is
  bounded and $|\hat \lambda - \lambda^*| = O_p(n_2^{-1/2})$, which proves our
  result.
\end{proof}

\subsection{Simulation Study 1}

@@

$$
\begin{aligned}
X_{1i} &\stackrel{ind}{\sim} N(2, 1) \\
X_{2i} &\stackrel{ind}{\sim} Unif(0, 4) \\
X_{3i} &\stackrel{ind}{\sim} N(0, 1) \\
\varepsilon_i &\stackrel{ind}{\sim} N(0, 1) \\
Y_{i} &= 3 X_{1i} + 2 X_{2i} + \varepsilon_i \\
\pi_{1i} &= \min(\Phi_3(-x_{3i} - 2), 0.7) \\
\pi_{2i|1} &= \min(\Phi_3(-x_{1i} + 1), 0.9).
\end{aligned}
$$

where $\Phi_3$ is the CDF of a t-distribution with 3 degrees of freedom.
%\textcolor{red}{The simulation setup is somewhat strange. The first order
%inclusion probability for the second-phase sampling, $\pi_{2i \mid 1}$, is a
%function of $x_{4i}$ which is completely independent of $(x_{1i}, x_{2i}, y_i)$.
%In this case, your second-phase sampling is essentially equivalent to the simple
%random sampling. }
This is a two-phase extension of the setup in \cite{kwon2024debiased}. We
consider a finite population of size $N = 10,000$ with both the Phase 1 and
Phase 2 sampling occuring under Poisson (Bernoulli) sampling. This yields a
Phase 1 sample
size of $E[n_1] \approx 1100$ and a Phase 2 sample size of
$E[n_2] \approx 300$. In the Phase 1 sample, we observe 
$(X_1, X_2)$ while in the Phase 2 sample we observe $(X_1, X_2, Y)$. This
simulation does not deal with model misspecification, and we compare the
proposed method for the parameter $\bar Y_N$ with four approaches:

\begin{itemize}
  \item[1.] $\pi^*$-estimator: $\hat Y_{\pi^*} = N^{-1} \sum_{i \in A_2}
    \frac{y_i}{\pi_{1i} \pi_{2i|1}},$
  \item[2.] Two Phase Regression estimator (TP-Reg): 
    $\hat Y_{reg} = \sum_{i \in A_1} \frac{\bf x_i' \hat \beta}{\pi_{1i}} + 
    \sum_{i \in A_2} \frac{1}{\pi_{1i}\pi_{2i|1}}(y_i - \bf x_i' \hat \beta)$ 
    where $\hat \beta = 
    \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$
    and $\bf x_i = (x_{1i}, x_{2i})^T$,
    % \textcolor{red}{Dr. Kim, should I modify the simulation so that the
    % regression estimator also includes $g(\pi_{2i|1}^{-1})$ as a covariate?}
  \item[3.] Debiased Calibration with Population Constraints (DC-Pop): This 
    solves

  \begin{equation}
    \argmin_{w_{2|1}} \sum_{i \in A_2} \frac{1}{\pi_{1i}} G(w_{2i})
    \text{ such that}
    \sum_{i \in A_2} w_{1i} w_{2i|1} z_i = \sum_{i \in U} z_i.
  \end{equation}

  \item[4.] Debiased Calibration with Estimated Population Constraints (DC-Est):
    This solves Equation~\eqref{eq:primal} with $q_i = 1$.
\end{itemize}

In addition to estimating the mean parameter $\bar Y_N$, we also construct
variance estimates $\hat V(\hat Y)$ for each estimate $\hat Y$. For each
approach we give the variance estimate in Table~\ref{tab:varforms}.

\begin{table}[ht!]
  \centering
  \begin{tabular}{lcc}
    \toprule
    Estimator & Estimated Variance & Notes \\
    \midrule
    $\pi^*$ & $N^{-2} \sum_{i \in A_2} \left(\pi_{2i|1}^{-2} - \pi_{2i|1}^{-1}\right)
    y_i^2 $ & \\
    TP-Reg  & 
    {\scriptsize $N^{-2}\left(\sum\limits_{i \in A_1} 
        \left(\frac{1 - \pi_{1i}}{\pi_{1i}^{2}}\right) \eta_i^2 
        + \sum\limits_{i \in A_2} \frac{1}{\pi_{1i} \pi_{2i|1}}(d_{2i|1} - 1)
        (y_i - \bf x_i' \hat \beta)^2 \right)$}
    & 
      {\scriptsize$\eta_i = \bf x_i \hat \beta + \frac{\delta_{2i}}{\pi_{2i|1}}(y_i - 
      \bf x_i \hat \beta)$} \\
      %$\hat \beta = \left(\sum_{i \in A_2} \bf x_i \bf x_i'\right)^{-1} \sum_{i \in A_2} \bf x_i y_i$ \\
    DC-Pop  & {\scriptsize$\sum\limits_{i \in A_2} c_i (y_i - q_i z_i \hat \gamma)^2 
              (w_{1i} \hat w_{2i|1}(\hat \lambda) / N)^2$}
            & {\scriptsize$c_i = ((1 - \pi_{1i}) \pi_{2i|1} + (1 - \pi_{2i|1}))$} \\
      DC-Est  & 
      \scriptsize{$N^{-2}\left(\sum\limits_{i \in A_1} \left(\frac{1 -
        \pi_{1i}}{\pi_{1i}^2}\right) \eta_i^2 + \sum\limits_{i \in A_2}
          \left(\frac{1}{\pi_{1i}\pi_{2i|1}}\right) (d_{2i|1} - 1) (y_i - q_i
          z_i \hat \phi)^2\right)$} & 
      {\scriptsize $\eta_i = q_i z_i \hat \phi +
          \frac{\delta_{2i|1}}{\pi_{2i|1}}(y_i - q_i z_i \hat \phi)$} \\
    \bottomrule
  \end{tabular}
  \caption{This table gives the formulas for each variance estimator used in
  Simulation 1. For the derivation of the estimated variance of the DC-Pop
estimatator see Appendix A.}
  \label{tab:varforms}
\end{table}

\textcolor{red}{Dr. Kim, please see Appendix A for my derivation of the variance
DC-Pop.}

%    Pi_dc <- 
%      diag((1 - p2_df$pi1) * p2_df$pi2 + (1 - p2_df$pi2)) * 
%      (w1i * dc_w / nrow(pop_df))^2
%    err <- p2_df$Y - t(gam_dc) %*% t(z_dc) * qi
%    vhat <- (err) %*% Pi_dc %*% t(err)

We run the simulation $1000$ times for each of these methods and compute the
Bias ($E[\hat Y] - \bar Y_N$), the RMSE ($\sqrt{\Var(\hat Y - \bar Y_N)}$), a 95\%
empirical confidence interval ($\sum_{b = 1}^{1000} |\hat Y^{(b)} - \bar Y_N| \leq 
\Phi(0.975)\sqrt{\hat V(\hat Y^{(b)})^{(b)}}$), and a T-test that assesses the
unbiasedness of each estimator where $\hat Y^{(b)}$ is the result from the $b$th
simulation replicate. The results are in Table~\ref{tab:tpdc-mean}.

\begin{table}[ht!]
  \centering
\input{tables/tpdcsim_mean.tex}
% This table was generated from /src/explore/20240411-tpdcsim.qmd
\caption{This table shows the results of Simulation Study 1. It displays the
Bias, RMSE, empirical 95\% confidence interval, and a t-statistic assessing the
unbiasedness of each estimator for the estimators: $\pi^*$, TP-Reg, DC-Pop, and
DC-Est.}
\label{tab:tpdc-mean}
\end{table}

@@

$$\hat X_1 = \sum_{i \in A_1}^{n_1} d_{1i} x_i \text{ and } 
\hat X_2 = \sum_{i \in A_2}^{n_2} d_{2i} x_i $$

where $d_{1i} = \pi_{1i}^{-1}$, $d_{2i} = \pi_{2i}^{-1}$, $\pi_{1i}$ is the
probability of $i \in A_1$ and $\pi_{2i} = \Pr(i \in A_2)$. 
%\textcolor{red}{I don't know if it makes sense to change the Phase 2
%selection probability as $\pi_{2i|1}$. (\textcolor{blue}{No. We do not know
%$\pi_{2i |1}$ and it is not necessary.}) Should I keep it at $\pi_{2i}$?
%\textcolor{blue}{(Yes, $\pi_{2i}$ is available to us.) } But if
%I do, how is the two phase setup useful?} 
%\textcolor{blue}{Non-nested two-phase sampling is different from the classical
%  two-phase sampling. We are not using the two-phase setup of Section 2  here.
%  We do not have to call it two-phase sampling. However, if we define $A_c=A_1
%  \cup A_2$ then $A_c$ serves the role of the first-phase sample in the
%  two-phase sampling. Then, our goal is to compute an estimator of $X$ from
%  $A_c$. The second-phase sample would be $A_2$. 
%}
We can combine these estimates to get
$\hat X_c = W \hat X_1 + (1 - W)\hat X_2$ for some matrix $W$ (see
\cite{merkouris2004combining} for the optimal choice of $W$), and can then define
a regression estimator as

$$
\hat Y_{NN, reg} = \hat Y_2 + (\hat X_c - \hat X_2)^T \hat \beta_q = 
\hat Y_2 + (\hat X_1 - \hat X_2)^T W^T\hat \beta_q 
$$

where % \textcolor{red}{(we may need transpose in $W$)}

$$\hat \beta_q = \left(\sum_{i \in A_2} \frac{x_i x_i^T}{q_i}\right)^{-1}
\sum_{i \in A_2} \frac{x_i y_i}{q_i} \text{ and }\hat Y_2 = \sum_{i \in A_2}
d_{2i} y_i. $$

@@

The methodology for the non-nested two phase sample is very similar to the setup
described as part of Topic 1. Given a strictly convex differentiable function,
$G: \mathcal{V} \to \mathbb{R}$, the goal is to solve

\begin{equation}\label{eq:nnopt}
\hat w_2 = \argmin_w \sum_{i \in A_2} G(w_{2i}) q_i \text{ such that } 
\sum_{i \in A_2} w_{2i} x_i = \hat X_{c} \text{ and } 
\sum_{i \in A_2} w_{2i} g(d_{2i}) q_i = \sum_{i \in U} g(d_{2i}) q_i
\end{equation}

for $g(x) = G'(x)$ and a known choice of $q_i \in \mathbb{R}$. 
%\textcolor{red}{(Instead of using $\hat{X}_{\rm GLS}$, we used $\hat{X}_c$
%before. Please make notation consistent. )}

The difference
between solving Equation~\ref{eq:nnopt} and Equation~\ref{eq:primal} is that the
estimator $\hat X_{c}$ is estimated from the combined sample $A_c= A_1 \cup A_2$. Before using
$\hat X_{c}$ in the debiased calibration estimator, we need to estimate it
from the non-nested samples.

We can get multiple estimates of $\hat X$,

$$\hat X_1 = \sum_{i \in A_1} d_{1i} x_i \text{ and }
\hat X_2 = \sum_{i \in A_2} d_{2i} x_i.$$

Let $V_1$ and $V_2$ be known variance matrices for $\hat X_1$ and $\hat X_2$,
then the optimal combined estimate is

$$\hat X_c = \frac{V_1^{-1} \hat X_1 + V_2^{-1} \hat X_2}{V_1^{-1} + V_2^{-1}}.$$

We can constuct a non-nested two phase estimator $\hat Y_{NNE}$ for $Y_N$ where
$\hat Y_{NNE} = \sum_{i \in A_2} \hat w_{2i} y_i$ and $\hat w_{2i}$ solves
Equation~\ref{eq:nnopt}. Like the classical two phase approach, to solve this
setup we minimize the Lagrangian,

\begin{equation}\label{eq:legragedc2}
  L(w_{2i}, \lambda) = \sum_{i \in A_2} G(w_{2i}) q_i + \lambda 
  \left( \hat T - \sum_{i \in A_2} w_{2i} z_i q_i\right).
\end{equation}

with

$$\hat T = 
\begin{bmatrix}
  \hat X_c \\ \sum_{i \in U} g(d_{2i}) q_i
\end{bmatrix}.
$$

\textcolor{red}{I am a little worried about the second term of $\hat{T}$. It is somewhat weird to assume that $T_g= \sum_{i \in U} g(d_{2i}) q_i$ is known. One simple way is to use a HT estimator  of $T_g$ from sample $A_2$. That is, use $\hat{T}_{g, {\rm HT}} = \sum_{i \in A_2} d_{2i} g( d_{2i}) q_i$ or even use a GREG estimator of $T_g$. 
%I wonder we can assume that an unbiased estimator of $T_g$ is available from $A_c=A_1 \cup A_2$. That is, let's assume that we have $\hat{T}_c = W \hat{T}_1 + (1- W) \hat{T}_2$ as the estimator  for $T$. 
}

Differntiating with respect to $w_{2i}$ and setting this expression equal to
zero, yields the fact that $\hat w_{2i}$ satisfies

$$ \hat w_{2i}(\hat \lambda) = g^{-1}(\hat \lambda^T z_i) $$

where $\hat \lambda$ is the solution to
